---
title: "Text and sentiment analysis"
description: "Performing sentiment analysis on books."
date: 03-31-2024
---

# Statistics on words

Let's first load the packages necessary for the analysis.

```{r message=FALSE, warning=FALSE}
#| label: Dependencies
# Charging dependencies
library(gutenbergr)
library(tidyverse)
library(tidytext)
library(stopwords)
library(hunspell)
library(SnowballC)
library(ggpattern)
library(ggwordcloud)
library(RColorBrewer)
library(textdata)
```

```{r message=FALSE, warning=FALSE}
#| label: Cleaning
wizard_of_oz <- gutenberg_download(54) |> select(text)

oz_df <- wizard_of_oz |>
  stringr::str_squish() |>
  tibble::as_tibble() |>
  tidytext::unnest_tokens(input = "value", 
                          output = "word", 
                          token = "words",
                          to_lower=TRUE) |>
  dplyr::anti_join(stopwords::stopwords(language = "en") |>
                     as_tibble(),
                   by=c("word"="value")) |>
  dplyr::mutate(stem = SnowballC::wordStem(word)) 
```

```{r}
#| label: Bar graph
bar_graph <- oz_df |>
  dplyr::group_by(word) |>
  dplyr::mutate(word_count = n()) |>
  dplyr::ungroup() |>
  dplyr::arrange(desc(word_count)) |>
  dplyr::distinct(word, .keep_all = TRUE) |>
  dplyr::slice_max(order_by = word_count, n=30) |>
  ggplot(aes(x=word |> reorder(word_count),y=word_count, fill=word_count))+ 
  ggpattern::geom_col_pattern(
    aes(pattern_fill = word_count),
    pattern = 'none',
    fill="seagreen4",
    show.legend = FALSE)+
  geom_label(aes(label=word_count), size = 3, color="white",hjust=-0.5)+
  coord_flip()+ 
  ylim(c(0,2500))+ 
  labs(x=NULL,y=NULL, 
       title="Words in The Jungle Book",
       subtitle = "30 most frequent words.")

bar_graph
```

Let's turn it into a wordcloud.

```{r}
#| label: Wordcloud
wordcloud <- oz_df |>
  dplyr::group_by(word) |> 
  dplyr::summarise(word_count = n()) |> 
  dplyr::distinct(word, .keep_all = TRUE) |> 
  dplyr::slice_max(order_by=word_count, n = 100) |>
  ggplot()+ 
  ggwordcloud::geom_text_wordcloud(aes(label = word, size = word_count, color=word_count)) +
  scale_size_area(max_size = 17) +
  scale_color_distiller(palette="RdYlGn", direction=-1)+
  theme_minimal()

wordcloud
```

# Sentiment analysis

::: callout-warning
The dataset bing from `tidytext` classifies the words envious, enviousness and enviously twice: once as positive, once as negative... Error ?
:::

```{r message=FALSE, warning=FALSE}
sentiment1 <- oz_df |> 
  select(-stem) |>
  dplyr::left_join(get_sentiments("bing"),
                   by=c("word")) |>
  dplyr::filter(!is.na(sentiment)) |>
  dplyr::group_by(sentiment) |>
  dplyr::summarise(sentiment_count = n()) |>
  # Plot
  ggplot(aes(x=sentiment,y=sentiment_count,fill=sentiment))+
  geom_col(show.legend = FALSE)+
  coord_flip()+
  labs(x=NULL,y="Word count",
       title = "The Wonderful Wizard of Oz, of L. Frank Baum", 
       subtitle = "Text analysis using Bing")+
  scale_fill_manual(values = c("#FFD0EC","#1F2544"))

sentiment1

```

```{r message=FALSE, warning=FALSE}
nrc <- tidytext::get_sentiments("nrc")

sentiment2 <- oz_df |>
  dplyr::left_join(nrc,
                   by=c("word"="word"), multiple="all") |>
  dplyr::filter(!is.na(sentiment)) |>
  dplyr::group_by(sentiment) |>
  dplyr::summarize(sentiment_count = n()) |>
  # Plot
  ggplot(aes(x=reorder(sentiment, sentiment_count), y=sentiment_count,fill=sentiment))+
  geom_col()+
  coord_flip()+
  labs(x=NULL,y="Word count",
       title = "Infinite Jest by David Foster Wallace", 
       subtitle = "Text analysis using NRC lexicon")+
  theme(legend.position = "none")+
  scale_fill_manual(values=c(
    "positive"="forestgreen",
    "negative"="orangered",
    "trust"="royalblue",
    "anticipation"="gold",
    "fear"="red",
    "sadness"="purple",
    "joy"="green",
    "anger"="black",
    "disgust"="brown",
    "surprise"="cyan")
    )

```

::: callout-warning
Why can't I use `tidytext::get_sentiment("nrc")` inside the left_join ?
:::

# Sentiment analysis of Pride and Prejudice, by Jane Austen

Let's begin by loading the data which is available in a R package (lucky me).

```{r message=FALSE, warning=FALSE}
#| label: Import
library("janeaustenr")
library("viridis")
pride <- janeaustenr::prideprejudice

```

Then we clean the data by tokenizing it and removing stop words.

```{r }
#| label: Text cleaning
pride_df <- pride |>
  stringr::str_squish() |>
  tibble::as_tibble() |>
  tidytext::unnest_tokens(input = "value", 
                          output = "text", 
                          token = "words", 
                          to_lower = TRUE) |>
  anti_join(stopwords::stopwords(language = "en") |>
                     as_tibble(), by=c("text"="value"))

```

We then make a nice use of a loop to get every different type of sentiment analysis available from the `tidytext::get_sentiment()` function.

```{r warning=FALSE}
#| label: Loop for SA
for (i in c("bing", "loughran", "nrc")){
  pride_df |>
    left_join(tidytext::get_sentiments(i), by=c("text"="word"), multiple="all") %>%
    filter(!is.na(sentiment)) |>
    group_by(sentiment) |>
    summarize(sentiment_count = n()) %>%
    {assign(paste0("sentiment_", i), ., envir = .GlobalEnv)}
}


```

::: callout-caution
This is a nice example of one way that the two pipes differ. Using the base R pipe ( \|\> ) doesn't allow for the use of brackets ( {} ), whereas `magrittr`'s pipe does ( %\>% ).
:::

AFINN is a bit different, variables names are different too, so we do not include it into the loop.

```{r warning=FALSE}
#| label: AFINN
sentiment_afinn <- pride_df |>
  left_join(tidytext::get_sentiments("afinn"), by=c("text"="word"), multiple="all") %>%
  filter(!is.na(value)) |>
  group_by(value) |>
  dplyr::mutate(n = n(),
                p = n/sum(n),
                c = case_when(value<0 ~ "Negative",
                              value==0 ~ "Neutral",
                              value>0 ~ "Positive") |>
                  factor())
```

## Some personnalisation

Let's make our own theme so that the graph are easier to customize.

```{r}
#| label: Create theme

theme_axel <- function () {
  theme(
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.grid.major.x = element_line(colour="black", linetype = "dashed"),
    panel.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4"),
    panel.border = element_rect(linetype = 1, fill=NA), 
    plot.background = element_rect(fill = "#fbf9f4", color = "#fbf9f4")
  )
}

```

## Bing

```{r}
#| label: Bing

plot_bing <- sentiment_bing %>%
  ggplot(aes(x=sentiment,y=sentiment_count,fill=sentiment))+
  geom_col(show.legend = FALSE)+
  coord_flip()+
  labs(x=NULL,y="Word count",
       title = "Pride and Prejudice, Jane Austen", 
       subtitle = "Text analysis using Bing")+
  scale_fill_viridis(discrete = TRUE, option="turbo") +
  theme_axel()

plot_bing
```

## NRC

```{r}
#| label: NRC

plot_nrc <- sentiment_nrc %>%
  ggplot(aes(x=sentiment,y=sentiment_count,fill=sentiment))+
  geom_col(show.legend = FALSE)+
  coord_flip()+
  labs(x=NULL,y="Word count",
       title = "Pride and Prejudice, Jane Austen", 
       subtitle = "Text analysis using NRC")+
  scale_fill_viridis(discrete = TRUE, option="turbo") +
  theme_axel()

plot_nrc
```

## LOUGRHAN

```{r}
plot_loughran <- sentiment_loughran %>%
  ggplot(aes(x=sentiment,y=sentiment_count,fill=sentiment))+
  geom_col(show.legend = FALSE)+
  coord_flip()+
  labs(x=NULL,y="Word count",
       title = "Pride and Prejudice, Jane Austen", 
       subtitle = "Text analysis using LOUGHRAN")+
  scale_fill_viridis(discrete = TRUE, option="turbo") +
  theme_axel()

plot_loughran

```

## AFINN

```{r}

plot_afinn <- sentiment_afinn %>%
  ggplot(aes(x=value,y=n,color=c))+
  geom_col(show.legend = FALSE)+
  geom_vline(xintercept = 0)+
  scale_x_continuous(breaks=seq(from=-5,to=5,by=1))+
  annotate("text",x=-3.5,y=4e+06, label ="Negative", color = "#30123BFF")+
  annotate("text",x=0.1,y=4e+06, label ="Positive", color = "#7A0403FF")+
  labs(x=NULL,y="Word frequency",
       title = "Pride and Prejudice, Jane Austen", 
       subtitle = "Text analysis using Afinn")+
  scale_color_viridis(discrete = TRUE, option="turbo") +
  theme_axel()

plot_afinn

```

## Comparison

| Sentiment lexicon | Pros                                                      |                                                                                                 Cons |
|:------------------------|----------------------|------------------------:|
| Bing              | Easy to understand                                        |                                                                                  Limited information |
| NRC               | Extensive information (sentiments AND emotions)           |                                                                                        to be found ? |
| LOUGHRAN          | New insights and nuances on what is expressed in the text | Created for use with financial documents, maybe not the most appropriate lexicon for analyzing books |
| AFINN             | Allows intensity analysis                                 |                                                                                  Limited information |

: Table 1: Comparison of sentiment lexicon available in tidytext.

# Function to clean text from `gutenbergr`

```{r}
#| code-fold: false

#Use id=54 as default
get_text <- function(id_gutenberg=54, lang="en") {
  gutenberg_download(id_gutenberg) |> 
    select(text) |>
    stringr::str_squish() |>
    tibble::as_tibble() |>
    tidytext::unnest_tokens(input = "value", 
                          output = "word", 
                          token = "words",
                          to_lower=TRUE) |>
  dplyr::anti_join(stopwords::stopwords(language = lang) |>
                     as_tibble(),
                   by=c("word"="value"))
}
```

# Function performing SA for any book from `gutenbergr`

```{r}
#| code-fold: false
sa <- function(book, lexicon=c("nrc", "bing", "loughran")){
  book %>%
    left_join(tidytext::get_sentiments(lexicon), by=c("word"="word"), multiple="all") %>%
    filter(!is.na(sentiment)) |>
    group_by(sentiment) |>
    summarize(sentiment_count = n()) %>%
    ggplot(aes(x=sentiment,y=sentiment_count,fill=sentiment))+
    geom_col(show.legend = FALSE)+
    coord_flip()+
    labs(x=NULL,y="Word count",
       subtitle = paste0("Text analysis using ", lexicon) )+
    scale_fill_viridis(discrete = TRUE, option="turbo") +
    theme_axel()
}
```
