---
title: "Web scrapping"
description: "Using web scrapping techniques to gather data from the web"
date: 05-04-2024
---

Let's start by installing the required dependencies.

```{r message=FALSE, warning=FALSE}
#| label: Loading packages
library(tidyverse)
library(rvest)
library(rlang)
library(reactable)
```

# 1. Get a list of someone's papers from Google scholar

Since there is no API, we define a function to retrieve any author's ID.

```{r}
#| label: Creating function to retrieve author's id
#| code-fold: show
get_author_id <- function(name){
  # add some warnings
  if(typeof(name)!="character"){stop("Please enter a character vector")}
  
  url <- paste0("https://scholar.google.com/citations?view_op=search_authors&mauthors=", paste(gsub(" ", "+", trimws(name)), sep="+"))
  
  id <- rvest::read_html(url) %>%
    rvest::html_elements(".gs_ai_name a") %>%
    rvest::html_attr("href") %>%
    stringr::str_extract("user=([^&]+)")%>% #extract id using some regex 
    str_replace("user=", "") %>% #clean the string
    first()
  
  #yet another warning
  if(rlang::is_empty(id) | is.na(id)){
    stop("No author was found")
  }else{
    return(id)
  }
}
```

Now we define a function to scrap 20 of the papers from this author.

```{r}
#| label: Function to retrieve the author's papers
#| code-fold: show
get_author_papers <- function(name, by=c("recent", "citations")){
  url <- paste0("https://scholar.google.com/citations?hl=fr&user=", get_author_id(name), "&view_op=list_works")
  
  #add an option for selecting paper based on recent or number of citation
  if(by=="recent"){url <- paste0(url, "&sortby=pubdate")}
  
  title <- url %>% 
    rvest::read_html() %>%
    rvest::html_elements(".gsc_a_at")%>%
    rvest::html_text() %>%
    tibble::as_tibble() %>%
    rename(title=value)
  
  year <- url %>% 
    rvest::read_html() %>%
    rvest::html_elements(".gsc_a_hc")%>%
    rvest::html_text() %>%
    tibble::as_tibble() %>%
    rename(year=value)
  
  #We need to do a bit of cleaning to remove the author's name from coauthors.
  #the strategy is the following: find the name of the author in the string and remove anything between the two commas where the name is.
  
  surname <- str_split(name, " ") %>% unlist() 
  surname <- surname[[length(surname)]]
  coauth <- 
    url %>% 
    rvest::read_html() %>%
    rvest::html_elements(".gsc_a_at+ .gs_gray")%>%
    rvest::html_text() %>% 
    #remove the part of the string containing the last name of the author 
    #and everything before and after in between commas
    str_replace(paste0("(?i)(?<=,|^)[^,]*", surname, "[^,]*(?=,|$)"), "") %>%
    #remove leading or trailing commas
    str_replace("^[,\\s]+|[,\\s]+$", "")%>%
    #replace two commas following each other by only one
    str_replace(",,", ",")%>% 
    tibble::as_tibble() %>%
    rename(coauth=value) 
  
  return(tibble(title, year, coauth))
}
```

::: {.callout-caution collapse="true"}
A small drawback of my method to get rid of the author's name among coauthors is that if two persons with the same last name publish a paper together (eg: a couple), then one of them will be removed (first occurrence of the last name). Although this can be a problem, it is unlikely to happen very often, and I have no other solution to offer as I cannot perform matching on the whole author's name, as first and middle names are usually abbreviated in different ways.
:::

Let's display those information for Daron Acemoglu.

```{r}
#| label: Apply for Daron Acemoglu

get_author_papers("Daron Acemoglu", by="recent") %>%
  reactable(
    searchable = TRUE, 
    striped = TRUE, 
    resizable = TRUE
    )
```

```{r include=FALSE}
status <- rvest::read_html("https://r4dev.netlify.app/sessions_workshop/06-scrap/06-scrap") %>%
  rvest::html_elements("#part-ii-beyond-static-websites p")%>%
  rvest::html_text()
```

::: callout-note
## What's next

Hopefully one day I'll learn how to interact with JS components to load more than 20 papers. For this to happen, I need Nelson's blog to be completed. The current status of the workshop is: `{r} print(status)`
:::

# 2. Scrap Wikipedia data

Load dependencies

```{r include=FALSE}
#| label: Loading dependencies
library(tidyverse)
library(rvest)
library(sf)
library(rnaturalearth)
library(ggthemes)
library(countrycode)
```

I've faced so many problems using rvest to try to scrap data from Wikipedia. For most pages, using the Selector gadget doesn't work, I've relied on the developer tools in Chrome and using xpath instead of css selectors...

I start by retrieving data on CO2 emissions on Wikipedia

```{r message=FALSE}
#| label: Retrieve data
co2_change <- rvest::read_html("https://en.wikipedia.org/wiki/List_of_countries_by_carbon_dioxide_emissions") %>%
  html_node(xpath='/html/body/div[2]/div/div[3]/main/div[3]/div[3]/div[1]/table') %>%
  html_table() %>%
  select(1,9) %>%
  rename(country=1, co2=2) %>%
  mutate(co2=as.numeric(gsub(",|%", "", co2))/100) %>%
  filter(!row_number() %in% c(1,210:213)) %>%
  mutate(co2_index=log(co2))

world_sf <- rnaturalearth::ne_download(returnclass = "sf")

```

We now create a map out of it.

```{r warning=FALSE}
#| label: Create map

co2_change %>%
  dplyr::mutate(iso3c = countrycode::countrycode(country, "country.name", "iso3c")) |>
  dplyr::right_join(world_sf |>
                      dplyr::select(ISO_A3_EH, geometry),
                    by = c("iso3c" = "ISO_A3_EH")) |>
  ggplot() +
  geom_sf(
    aes(fill = co2_index,
        geometry = geometry),
    alpha = 0.5) +
  ggplot2::scale_fill_gradientn(
    colours=c("darkgreen", "lightgreen", "yellow", "darkred"),
    values = scales::rescale(c(-2,0,0.001, 2.5)), 
    na.value = "white",
    limits = c(-2, 2.5),
    oob = scales::oob_squish, 
    breaks=c(-2,0,2.5),
    labels=c("-2 (decrease)", "0 (equal)", "2.5 (increase)"),
    name="Evolution (in log)"
  )+
  labs(
    title = "CO2 Emissions evolution since 1990",
    caption = "Source: Crippa et al (2023). GHG emissions of all world countries – 2023. Luxembourg: Publications Office of the European Union. Retrieved from Wikipedia. "
  ) +
  ggthemes::theme_map()
```
